The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-04 sandbox/note_km.tex
km note: update discussion of "true occ rate"

Consider the case when the $Z_i$'s are identical. If we compare 
$\Gamma_a(r_a)$ with $\Gamma(r)$, we must impose that $r_a$ and $r$ are 
indistinguishable (i.e.: the observer cannot tell them apart).
Note that the true rate density is
\begin{equation}
\Gamma(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right]
\end{equation}
We then get a correction factor
\begin{equation}
X_\Gamma \equiv \left. \frac{\Gamma_a(r_a)}{\Gamma(r)} 
\right|_{r_a\rightarrow r}
=
\frac{2^{\frac{\delta}{2}+2}\mu + 1}{\mu + 1}.
\end{equation}

For the case of ${\rm BF}=0.1$, $\mu\approx 0.153$. Taking $\delta=-2.92$ from 
Howard et al. (2012),  we get a correction factor $X_\Gamma = 
\Gamma_a/\Gamma = 1.06$.
In other words, the apparent rate density is an {\it over}estimate of the true 
rate density, with a relative error of 6\%.

@@

Note that the true rate density is
\begin{equation}
\Gamma(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right].
\end{equation}
The rate density correction factor, $X_\Gamma = \Gamma\a/\Gamma$, will be
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma_a(r_a)}{\Gamma(r)} 
\right|_{r_a\rightarrow r} \\
&=
\frac{1+2\mu}{3(1+\mu)}
\left[1 + \mu
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta}{2}+2} +
\int {\rm d}q\,q^{\beta+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\delta=-2.92$, the term in $(\ldots)\approx 
1.41425 \approx \sqrt{2}$, where the latter surprising numerical fluke holds 
to the first four decimal places.
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma = 1.28$, an apparent rate 
which is {\it higher} than the true rate by 28\%!

>>>>

Consider the case when the $Z_i$'s are identical. If we compare 
$\Gamma_a(r_a)$ with $\Gamma(r)$, we must impose that $r_a$ and $r$ are 
indistinguishable (i.e.: the observer cannot tell them apart).
Note that the true rate density for stars selected in the sample is
\begin{equation}
\Gamma_{\rm sel}(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right],
\end{equation}
while the true rate density for a volume-limited sample of stars is
\begin{equation}
\Gamma_{\rm vl}(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2(n_b/n_s)} + \frac{(n_b/n_s)}{1+2(n_b/n_s)} (Z_1+Z_2)
\right],
\end{equation}
for $n_b$ and $n_s$ the number density of binaries and singles in a 
volume-limited sample.

The stand-alone term ``true rate density'' is ill defined, and should probably 
be avoided.
In different contexts, it might refer to either a) 
$\Gamma_{\rm sel}(r)$, b) $\Gamma_{\rm vl}(r)$, or even c) $\Gamma_0(r)$, the 
rate around singles.
The meaning that most people probably {\it want} true rate density to have is
$\Gamma_0(r)$~---~simply ``I have a star that is single; how many planets are 
there?''.
This is also a sensible definition because it is {\it independent} of any 
complications of planet formation that would make $Z_1$ or $Z_2$ different 
from $Z_0$.
One way to deal with this, rather than making the arbitrary choice, would be 
to just plot out all the possible options.

We could define a ``correction factor'' relative to $\Gamma_0$, or to 
$\Gamma_{\rm sel}$, or to $\Gamma_{\rm vl}$.
If the $Z_i$'s are equal, these are all identical:
\begin{align}
\notag
X_\Gamma &\equiv 
\left. \frac{\Gamma_a(r_a)}{\Gamma_0(r)} 
\right|_{r_a\rightarrow r,\ Z_i{\rm 's}\ {\rm equal}} \\
\notag&=
\left. \frac{\Gamma_a(r_a)}{\Gamma_{\rm sel}(r)} 
\right|_{r_a\rightarrow r,\ Z_i{\rm 's}\ {\rm equal}} \\
\notag&=
\left. \frac{\Gamma_a(r_a)}{\Gamma_{\rm vl}(r)} 
\right|_{r_a\rightarrow r,\ Z_i{\rm 's}\ {\rm equal}} \\
&=
\frac{1 + 2^{\frac{\delta+3}{2}}\mu}{1 + \mu}.
\end{align}
For the case of ${\rm BF}=0.1$, $\mu\approx 0.153$. Taking $\delta=-2.92$ from 
Howard et al. (2012),  we get a correction factor $X_\Gamma = 
\Gamma_a/\Gamma = 1.004$.
In other words, the apparent rate density is an {\it over}estimate of the rate 
density of selected stars, with a relative error $\delta \Gamma = |\Gamma - 
\Gamma\a | / \Gamma$ of 0.4\%.

@@

Note that the true rate density for the selected stars is
\begin{equation}
\Gamma_{\rm sel}(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right].
\end{equation}
The correction factor with respect to selected stars, $X_\Gamma = 
\Gamma\a/\Gamma_{\rm sel}$, will be
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a)}{\Gamma_{\rm sel}(r)} 
\right|_{r_a\rightarrow r,\ Z_i{\rm 's\ equal}} \\
&=
\frac{1+2\mu}{3(1+\mu)}
\left[1 + \mu
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta}{2}+2} +
\int {\rm d}q\,q^{\beta+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\delta=-2.92$, the term in $(\ldots)\approx 
1.41425 \approx \sqrt{2}$, where the latter surprising numerical fluke holds 
to the first four decimal places.
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_{\rm sel} = 1.28$, an 
apparent rate which is {\it higher} than the true rate by 28\%!
Note that this result is the same as $\Gamma\a/\Gamma0$ and 
$\Gamma\a/\Gamma_{\rm vl}$ in the limit where all the $Z_i$'s are equal.
